Stochastic Wind Power Forecasting

### 2020-04-17 LaTeX_Code/Paper/paper_1line_Renzo.tex
Added initial guess in Paper. Still not complete.

\usepackage{graphicx}
\usepackage{bm}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{todonotes}
%\usepackage[normalem]{ulem}
\usepackage{float}
\usepackage{commath}
\usepackage{fancybox}
\usepackage{url}

\theoremstyle{definition}
\newtheorem{Def}{Definition}[section]
\newtheorem{Eg}{Example}[section]
\newtheorem{Rm}{Remark}[section]
\newtheorem{Ex}{Exercise}[section]
\newtheorem{Prop}[Def]{Proposition}
\newtheorem{Lem}[Def]{Lemma}
\newtheorem{Thm}[Def]{Theorem}
\newtheorem{Cor}[Def]{Corollary}
\newtheorem{Rem}[Def]{Remark}
\newtheorem{Ass}[Def]{Assumption}
\newtheorem{Met}[Def]{Method}
\newtheorem{Res}[Def]{Result}
\newcommand{\red}{\textcolor{red}}

@@

\subsection{Optimization} \label{opt_sec}
{\color{red} Rewrite the discussion to obtain fast initial estimate for the two unknow parameters.}

To initialize the optimization process for the likelihood function of the process $V$, we solve the following least-squares problem which gives us first estimates of the mean reversion parameter $\theta_0$: {\color{red} Add explanation regarding the motivation for this objective function.}

\begin{equation}
\theta_0\approx\arg\min_{\theta} \sum\limits_{j}^M \sum\limits_i^N \left( v_{i+1,j}  - v_{i,j}- \left( - \theta_t v_{i,j}\right) \left(t_{i+1,j} - t_{i,j} \right)  \right)^2,
\label{inital_theta}
\end{equation}

 where $v_{i,j}=x_{i,j}-p_{i,j}$,  $x_{i,j}$ is historical wind power production and $p_{i,j}$ is the wind power forecast.

 By assuming ergodicity {\color{red} not ergodicity, instead simply for the \textbf{quadratic variation formula}.}, we can obtain a first estimate on the product of the parameters as follows,

\begin{equation}
\theta_0 \alpha \approx \frac{1}{2M\Delta t} \sum\limits_j^M \frac{ \sum\limits_i^N (x_{i+1,j}  - x_{i,j})^2}{ \sum\limits_i^N x_{i,j}(1-x_{i,j}) } \,.
\end{equation}

Solving for $\alpha$, we have both first estimates and we can also estimate $\delta$ solving the problem

\begin{equation}
\delta\approx\arg\min_{\delta}\mathcal{L}_{\delta}(\bm{\theta},\delta; V^{M,1}) = \arg\min_{\delta}\prod\limits_{j=1}^M \rho_0 (V_{j, t_0}|V_{j, t_{-\delta}};\bm{\theta},\delta) \,.
\label{likelihood_delta}
\end{equation}

With our estimations of $\bm{\theta}$, we can start the inference processes as follows:
\begin{enumerate}
\item[Step 1.]  initialize;
\item[Step 2.]  optimize the log-likelihood function using  Nelder-Mead optimization algorithm on a mini-batch sampled with replacement;
\item[Step 3.] check if  accuracy threshold is reached. Else, re-initialize the optimization with the most recent result on a larger mini-batch.
\end{enumerate}
%---END SECTION 4---

>>>>

\usepackage{graphicx}
\usepackage{bm}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{todonotes}
%\usepackage[normalem]{ulem}
\usepackage{float}
\usepackage{commath}
\usepackage{fancybox}
\usepackage{url}
\usepackage[margin=0.5in]{geometry}

\theoremstyle{definition}
\newtheorem{Def}{Definition}[section]
\newtheorem{Eg}{Example}[section]
\newtheorem{Rm}{Remark}[section]
\newtheorem{Ex}{Exercise}[section]
\newtheorem{Prop}[Def]{Proposition}
\newtheorem{Lem}[Def]{Lemma}
\newtheorem{Thm}[Def]{Theorem}
\newtheorem{Cor}[Def]{Corollary}
\newtheorem{Rem}[Def]{Remark}
\newtheorem{Ass}[Def]{Assumption}
\newtheorem{Met}[Def]{Method}
\newtheorem{Res}[Def]{Result}
\newcommand{\red}{\textcolor{red}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}

@@

\subsection{Optimization} \label{opt_sec}

\subsubsection{Initial guess}

To guarantee the good behave for our optimization algorithm, we aim to start the optimization as close as we can from the optimal parameters. We use \textbf{Lease Square Minimization} and \textbf{Quadratic Variation} to find an initial guess from the data.

\begin{itemize}

\item \textbf{Least Square Minimization:} We consider the transition $\Delta V_i=V_{i+1}-V_i$ with $\Delta t=t_{i+1}-t_i$. $(V_{i+1}|V_i)$ is a random variable which conditional mean can be approximated by the solution of the system
\begin{equation*}
\begin{cases}
\dif \E\left[V\right]&=-\theta_t^\epsilon\E\left[V\right]\dif t\\
\E\left[V(t_i)\right]&=V_i,
\end{cases}
\end{equation*}
evaluated in $t_{i+1}$ (i.e., $\E\left[V(t_{i+1})\right]$). Then, the random variable $(V_{i+1}-\E\left[V(t_{i+1})\right])$ has approximately zero mean.\\
If we assume that $\theta_t^\epsilon={\color{red}c}\in\R^+$ for all $t\in[t_i,t_{i+1}]$, then $\E\left[V(t_{i+1})\right]=V_ie^{-{\color{red}c}\Delta t}$. If we have a total of $n$ transitions, we can write the regression problem for the conditional mean with $L^2$ loss function as
\begin{equation}
{\color{red}c}^*\approx\arg\min_{{\color{red}c}\geq0}\left[\sum_{i=1}^n\left(V_{i+1}-\E\left[V(t_{i+1})\right]\right)^2\right]=\arg\min_{{\color{red}c}\geq0}\left[\sum_{i=1}^n\left(V_{i+1}-V_ie^{-{\color{red}c}\Delta t}\right)^2\right].
\label{Eq-1}
\end{equation}
\begin{equation*}
\begin{split}
\frac{\partial f}{\partial{\color{red}c}}&=\sum_{i=1}^n2(-V_i)(-\Delta t)(V_{i+1}-V_i(1-\theta_0\Delta t))\\
&=\sum_{i=1}^n2V_i\Delta t(V_{i+1}-V_i(1-{\color{red}c}\Delta t))\\
&=\sum_{i=1}^n2V_{i+1}V_i\Delta t-2V_i^2\Delta t+2V_i^2\Delta t^2{\color{red}c}.
\end{split}
\end{equation*}
Then, ${\color{red}c}^{**}$ satisfies
\begin{equation}
{\color{red}c}^{**}\approx\frac{\sum_{i=1}^nV_i(V_i-V_{i+1})}{\Delta t\cdot\sum_{i=1}^n(V_i)^2}.
\label{Eq-4}
\end{equation}
Notice that ${\color{red}c}^{**}$ has dimension $\mathbf{time}^{-1}$.

\item \textbf{Quadratic Variation:} We approximate the SDE by its E-M scheme. In particular, we approximate the It\^o quadratic variation with the discrete one:
\begin{itemize}

\item It\^o process quadratic variation: $[V]_t=\int_0^t\sigma_s^2\dif s$.
\item Discrete process quadratic variation: $[V]_t=\sum_{0<s\leq t}(\Delta V_s)^2$.

\end{itemize}
\quad\\
\quad\\
Then, considering $\Delta t$ the time between measurements, we approximate:
\begin{equation}
\theta_0^*\alpha^*\approx\frac{\sum_{i=1}^n(\Delta V_i)^2}{2\Delta t\sum_{i=1}^n(V_i+p_i)(1-V_i-p_i)}.
\label{Eq-2}
\end{equation}

In slide ({\color{blue}\ref{theta_t}}) we defined the parameter $\epsilon$ which has as only condition $0<\epsilon<\frac{1}{2}$. We want to estimate the set of parameters $(\theta_0,\alpha,\epsilon)$, we call to the estimations $(\theta_0^*,\alpha^*,\epsilon^*)$.\\
\quad\\
Recall from slide ({\color{blue}\ref{S1}}) that, to use the LSM estimation, we assume $\theta_t^\epsilon={\color{red}c}\in\R^+$. Recall also our definition for $\theta_t^\epsilon$ (equation ({\color{blue}\ref{Eq-11}})):
\begin{equation*}
\theta_t^\epsilon=\max\left(\theta_0,\frac{\alpha\theta_0+2|\dot{p}_t^\epsilon|}{2\min(1-p_t^\epsilon,p_t^\epsilon)}\right).
\end{equation*}
Fixed $\epsilon$, we call $\mathcal{V}=\{\Delta V_i^\epsilon\}_{i=1}^n$ to the set of $n$ error transitions where for each measurement $X_i$, we have that $V_i=X_i-p_i^\epsilon$. $\mathcal{P}=\{p_i^\epsilon\}_{i=1}^n$ is the corresponding set of forecasts.\\
Now, if we also fix $\theta_0$ and $\alpha$, we can define the set of indexes $\mathbf{I}=\{i\in\{1,\dots,n\}:\text{ the LSM estimation will estimate }\theta_0\}$ and $\mathbf{J}=\left\{j\in\{1,\dots,n\}:\text{ the LSM estimation will estimate }\frac{\theta_0\alpha}{\epsilon}\right\}$.

From the definition of $\theta_t^\epsilon$: We have that for $\epsilon<<1$, and $p_t=\epsilon$ or $p_t=1-\epsilon$, the approximation $\theta^\epsilon_t\approx\frac{\theta_0\alpha}{\epsilon}$ holds. Then, for $\epsilon$ small enough, $\mathbf{J}$ can be approximated by $$\mathbf{J}\approx\tilde{\mathbf{J}}=\{j\in\{1,\dots,n\}:p_j^\epsilon\in\{\epsilon,1-\epsilon\}\}.$$
Now, coming back to the definition of $\theta_t^\epsilon$, we have that it is more likely that $\theta_t^\epsilon=\theta_0$ if $p_t^\epsilon\approx\frac{1}{2}$. Then, we can approximate $\mathbf{I}$ by $$\mathbf{I}\approx\tilde{\mathbf{I}}=\left\{i\in\{1,\dots,n\}:p_i\in(\gamma,1-\gamma)\right\},\ \gamma\approx\frac{1}{2},\ \gamma<\frac{1}{2}.$$

Recall the quadratic variation estimator ({\color{blue}\ref{Eq-2}}) from slide ({\color{blue}\ref{QV}}). It uses all the transitions $\mathcal{V}$ and estimated the product $\theta_0\alpha$.\\
\quad\\
We get ${\color{orange}\theta_0^*\alpha^*\approx0.10}$, from where we calculate ${\color{orange}\alpha^*\approx0.08}$.

As we have an approximated value for $\theta_0\alpha$, if we can estimate $\frac{\theta_0\alpha}{\epsilon}$, then we can estimate $\epsilon$. In slide ({\color{blue}\ref{Slide:Defs}}), we showed that for $\epsilon<<1$, the LSM estimation using indexes from $\tilde{\mathbf{J}}$ is an estimator for $\frac{\theta_0\alpha}{\epsilon}=:k$.\\
\quad\\
However, which values of $\epsilon$ satisfies $\epsilon<<1$? We choose an initial guess for $\epsilon$ (call it $\epsilon_0$), and iterating we aim to converge to some local minimum. We proceed with the following steps:

\begin{itemize}

\item We sample $\epsilon_0$ from $\mathcal{U}[0.01,0.1]$ and load $\epsilon\gets\epsilon_0$.
\item We create $\tilde{\mathbf{J}}$ and use the LSM estimation to find $k$. If $k<\theta_0^*$, then the assumption $\theta_t^\epsilon={\color{red}c}\in\R^+$ is wrong and we reduce the value of $\epsilon$, i.e., $\epsilon\gets\epsilon*0.999$. If $k\geq\theta_0^*$, we load $\epsilon\gets\frac{\theta_0^*\alpha^*}{k}$ (we allow a maximum relative change of 1\%). We repeat this step *100 times*.
\item We repeat steps 1 and 2 *500 times*. After, we plot the results.

%---END SECTION 4---
